Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-16 intro.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

The rest of the paper is organized as follows.
In Section~\ref{sec:svdf} we introduce our stochastic dynamics framework, recalling the single-track vehicle model with nonlinear tires used for planning, and formulate both the continuous-time planning problem and its discrete-time counterpart via direct-collocation. Here, we also derive the probabilistic constraint back-off formulation.
Section~\ref{sec:open_loop_planning} then outlines the open-loop horizon-based covariance-propagation planning approach, while Section~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner. In Section~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
Finally, Section~\ref{sec:conclusions} draws conclusions and offers directions for future work.

>>>>

The rest of the paper is organized as follows.
In Sec.~\ref{sec:svdf} we introduce our stochastic dynamics framework, recalling the single-track vehicle model with nonlinear tires used for planning, and formulate both the continuous-time planning problem and its discrete-time counterpart via direct-collocation. Here, we also derive the probabilistic constraint back-off formulation.
Sec.~\ref{sec:open_loop_planning} then outlines the open-loop horizon-based covariance-propagation planning approach, while Sec.~\ref{sec:closed_loop_planning} presents the closed-loop, covariance-aware planner. In Sec.~\ref{sec:performance_comparison} we compare both methods on a representative track scenario, reporting parameter-sensitivity studies, analyzing the influence of different sets of constraints, and performance trade-offs.
Finally, Sec.h~\ref{sec:conclusions} draws conclusions and offers directions for future work.

### 2025-06-16 open_loop_planning.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

\section{Open-loop planning via H-steps ahead predictions}
\label{sec:open_loop_planning}
In this section, we address the problem of planning under uncertainty for a stochastic system.
At each grid point $k$ of the discretized trajectory the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps.
In the planning problem, we conservatively evaluate the maximum propagation of the covariance, that is, the matrix evolved over $H$ steps, to ensure the tightest possible enforcement of constraints. This maximizes the robustness of the resulting solution, as the constraint back-off is evaluated in the worst-case scenario within the prediction horizon.

@@

This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Section~\ref{sec:discretization}. In particular, Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}. Equation~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation equations for all versions of the covariance matrix propagated from previous steps. To support the evolution of $H$ versions, we introduce $H \times d$ matrices $\bSi_k^j = (\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, representing the collocation values of each version $j$ of the covariance matrix within interval $k$.

>>>>

\section{Open-loop planning via H-steps ahead predictions}
c In this section, we address the problem of planning under uncertainty for a stochastic system.
At each grid point $k$ of the discretized trajectory the covariance matrix is propagated in a feed-forward fashion for a fixed prediction horizon of $H$ steps.
In the planning problem, we conservatively evaluate the maximum propagation of the covariance, that is, the matrix evolved over $H$ steps, to ensure the tightest possible enforcement of constraints. This maximizes the robustness of the resulting solution, as the constraint back-off is evaluated in the worst-case scenario within the prediction horizon.

@@

This approach requires modifications to the formulation~\eqref{eq:DOCP} introduced in Sec.~\ref{sec:discretization}. In particular, Eqs.~\eqref{eq:DOCPdP}, \eqref{eq:DOCPdPIC}, and \eqref{eq:DOCPconstraints} are replaced by Eqs.~\eqref{eq:DOCPdPopenloop}, \eqref{eq:DOCPdPinitopenloop}, and \eqref{eq:DOCPconstraintsopenloop}. Equation~\eqref{eq:DOCPdPopenloop} collects the continuity and collocation equations for all versions of the covariance matrix propagated from previous steps. To support the evolution of $H$ versions, we introduce $H \times d$ matrices $\bSi_k^j = (\bSi_{k,1}^j, \ldots, \bSi_{k,d}^j)$, representing the collocation values of each version $j$ of the covariance matrix within interval $k$.

### 2025-06-16 performance_comparison.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

 Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
\begin{align}
\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
\end{align}
These standard deviations have been chosen based on the typical range of variation of each state. To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.

@@

%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sections~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).

First we show results obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%---Delta key signals
The comparison between the nominal case and the three robustified cases is illustrated in Fig.~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- high, medium, and low --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

An analogous analysis has been conducted using the closed-loop approach described in Section~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
The same key signals displayed for the previous case are shown in Fig.~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Fig.~\ref{fig:cl_saturation}.
From one side, the results appear very similar to those obtained with the open-loop method. While the closed-loop method offers a more rigorous framework by enabling seamless covariance propagation over the entire lap, the observed agreement indicates that open-loop planning remains a valid and effective strategy, as long as the prediction horizon is appropriately selected.

>>>>

 Assuming the state vector is ordered as described in Sec.~\ref{sec:vehicle_model}, the baseline initial covariance matrix is set to $\bP_0=\diag(\bbsig^2)$, with
\begin{align}
\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T.
\end{align}
These standard deviations have been chosen based on the typical range of variation of each state. To provide a concrete interpretation at the starting condition, the assumption on first state $u \sim \calN(\mu_u,\sig_u^2)$, with $\sig_u=0.1$ m/s, implies that
$\Pr\{\mu_u-\sig_u \leq u \leq \mu_u+\sig_u\}\approx 68\%$ and $\Pr\{\mu_u-3\sig_u \leq u \leq \mu_u+3\sig_u\}\approx 99\%$.

@@

%It is of interest to analyze how a robustified constraint---or a combination of the two described in Sections~\ref{sec:FLC} and~\ref{sec:TLC}---influences the driving style.
We now investigate how the inclusion of different kinds of robustified constraints --- individually or in combination, as described in Sec.~\ref{sec:FLC} and~\ref{sec:TLC} --- influences the driving style.
We compare the nominal feed-forward optimal trajectory, obtained without robustified constraints, with those resulting from a robustified track limit constraint (denoted as TLC), a robustified friction limit constraint (denoted as FLC), and a scenario where both constraints are robustified (denoted as TLC+FLC).

First we show results obtained with the open-loop method described in Sec.~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%All the optimization are obtained with the open-loop method described in Section~\ref{sec:open_loop_planning}, setting $H=4$, and $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$, corresponding to a 90\% probability of satisfying both constraints.
%---Delta key signals
The comparison between the nominal case and the three robustified cases is illustrated in Fig.~\ref{fig:ol_telemetries}, in terms of the deviation of selected signals from to their nominal counterparts.
The panels display variations on the longitudinal speed $\De u$ (first panel), wheel steering angle $\De \de$ (second panel), total longitudinal force $\De F_x$ (third panel), and lateral deviation of the center of mass (CoM) from the track centerline $\De e$ (fourth panel).
Except for the $\De u$ panel, each plot includes dashed black lines that indicate the sign of the nominal signal.
These indicators are shown at three distinct vertical levels --- high, medium, and low --- corresponding respectively to positive, near-zero, and negative values of the nominal signal.

@@

An analogous analysis has been conducted using the closed-loop approach described in Sec.~\ref{sec:closed_loop_planning}, setting $\ga^\textrm{TLC}=\ga^\textrm{FLC}=1.28$.
The same key signals displayed for the previous case are shown in Fig.~\ref{fig:cl_telemetries}, while the axle saturation ratios $S_j$ are illustrated in Fig.~\ref{fig:cl_saturation}.
From one side, the results appear very similar to those obtained with the open-loop method. While the closed-loop method offers a more rigorous framework by enabling seamless covariance propagation over the entire lap, the observed agreement indicates that open-loop planning remains a valid and effective strategy, as long as the prediction horizon is appropriately selected.

### 2025-06-16 stochastic_vehicle_dynamics.tex
[MGa] abbrv. Sec. where necessary
come commas and some amendments to .bib

The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position $\left(x_G, y_G\right)$ and orientation $\psi$ of the barycentrical reference frame. As illustrated in Fig.~\ref{fig:vehicle_model} $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.

>>>>

The state vector $\bx$ has six components and includes the three kinematic quantities $u$, $v$, and $r$, along with the position $\left(x_G, y_G\right)$ and orientation $\psi$ of the barycentrical reference frame. As illustrated in Fig.~\ref{fig:vehicle_model}, $u$ and $v$ represent the longitudinal and lateral velocity of the CoM w.r.t its reference frame, while $r$ represents vehicle's yaw rate.
